Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-25 appendix.tex
Edits to appendix.  Time to move back to the test problem section, which I have been avoiding because it's very, very long.

In this appendix, we provide details for the various interpolation
methods available in the code.  We assume throughout that we are
dealing with Cartesian coordinates and cells of equal sizes, although
we allow for arbitrary an refinement factor $r$ between cells at
different levels.

@@

In two dimensions, the procedure is very similar in that we linearly
interpolate parent values to the four cell corners: $Q_{-1/2, -1/2}$,
$Q_{-1/2, 1/2}$, $Q_{1/2, -1/2}$, $Q_{1/2, 1/2}$ (by averaging parent
cell-centered parent values).  We then compute monotonic slopes across
the two diagonals, since for a linear function, extrema occur at
corners:

@@

Finally, we write down the three-dimensional version -- unfortunately,
here the number of monotonicity constraints is four (the 4 diagonals
across the 8 opposing corners of the cube), while the number of slopes
is three, so the problem is over constrained.  Somewhat arbitrarily,
we adopt the following procedure.  As before, we compute $\Delta Q_0$,
$\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$ with the minmod limiter
across the 4 diagonals.  We then define

@@

We also provide a variant on the above procedure, with two changes.
The first is that the slopes across the diagonals ($\Delta Q_0$, etc.)
are computed directly rather than with the minmod limiter
(e.g. $\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).  To
ensure positivity in the resulting interpolation when applied to
positive conserved quantities, the slopes are limited so that the
smallest corner value is 0.2 the cell center value.  The procedure
described above is then applied to turn these four slopes into a
linear interpolation.

@@

\begin{equation}
\tilde{Q_0} = \sum_{i=0}^{r-1} q_i
\end{equation}
is not, in general, equal to $Q_0$.  We can retain conservation by
adding the factor $(Q_0 - \tilde{Q_0})/r$ to the interpolated values
(or by multiplying the interpolated values by the ratio
$Q_0/\tilde{Q_0}$).  Unfortunately, the result of this procedure does
not preserve monotonicity -- it can introduce local minima and maxima
at parent cell boundaries.  We can then attempt to correct that by
taking a weighted average between the interpolated values on the
subgrid and the parent value with weights computed such that the
interpolated values at the edge of the parent cell are not local
maxima compared to the interpolated values in the neighboring parent
cell.

>>>>

In this appendix, we provide details for the various interpolation
methods available in the code.  We assume throughout that we are
dealing with Cartesian coordinates and cells of equal sizes, although
we allow for an arbitrary refinement factor $r$ between cells at
different levels.

@@

In two dimensions, the procedure is very similar in that we linearly
interpolate parent values to the four cell corners: $Q_{-1/2, -1/2}$,
$Q_{-1/2, 1/2}$, $Q_{1/2, -1/2}$, $Q_{1/2, 1/2}$ (by averaging parent
cell-centered parent values).  We then compute monotonic slopes across
the two diagonals, since for a linear function extrema occur at
corners:

@@

Finally, we write down the three-dimensional version -- unfortunately,
here the number of monotonicity constraints is four (the 4 diagonals
across the 8 opposing corners of the cube), while the number of slopes
is three, so the problem is over-constrained.  Somewhat arbitrarily,
we adopt the following procedure.  As before, we compute $\Delta Q_0$,
$\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$ with the minmod limiter
across the 4 diagonals.  We then define

@@

We also provide a variant on the above procedure, with two changes.
The first is that the slopes across the diagonals ($\Delta Q_0$, etc.)
are computed directly rather than with the minmod limiter
(e.g. $\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).  To
ensure positivity in the resulting interpolation when applied to
positive conserved quantities, the slopes are limited so that the
smallest corner value is 0.2 of the cell center value.  The procedure
described above is then applied to turn these four slopes into a
linear interpolation.

@@

\begin{equation}
\tilde{Q_0} = \sum_{i=0}^{r-1} q_i
\end{equation}
is not, in general, equal to $Q_0$.  We can retain conservation by
adding the factor $(Q_0 - \tilde{Q_0})/r$ to the interpolated values
(or by multiplying the interpolated values by the ratio
$Q_0/\tilde{Q_0}$).  Unfortunately, the result of this procedure does
not preserve monotonicity -- it can introduce local minima and maxima
at parent cell boundaries.  We can then attempt to correct that by
taking a weighted average between the interpolated values on the
subgrid and the parent value, with weights computed such that the
interpolated values at the edge of the parent cell are not local
maxima compared to the interpolated values in the neighboring parent
cell.
